---
title: "create_obj_sma_processed"
author: "Jian Wu"
date: "2025-09-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load packages

```{r }
library(Seurat)
library(STutility)
```
## load sma object data

obj.list.sma: object that is produced from multimodality integration, including both rna and msi modalities
obj.all.rna: object of rna that went through gene expression analysis and annotation

```{r pressure, echo=FALSE}

obj.list.sma <- readRDS("/Users/wujian/Documents/GitHub/stratums/R_objects/se.multi.list.new")

obj.all.rna <- readRDS("/Users/wujian/Documents/MARCO/R_objects/mPDStrRNA.10xFFBM.STUF.An")
obj.list.rna <- SplitObject(obj.all.rna, split.by = "Sample.ID")
```

## create and save a toy dataset (only one sample) to test the visualization effect

obj.list.sma: object that is produced from multimodality integration, including both rna and msi modalities
obj.all: object of rna that went through gene expression analysis and annotation

obj.list.sma and obj.all went through different filtering, resulting different remaining spot sets
obj.sub is the intersected set between the two (sma integration and gene expression pipeline)

the file size of obj is larger than that of obj.sub.

transfer lesion info from obj.all to obj.sub

```{r pressure, echo=FALSE}
obj.sma <- obj.list.sma[[5]]
obj.rna


obj.sub <- obj[,colnames(obj) %in% colnames(obj.all)]
obj.
obj.sub$lesion <- obj.all$lesion[colnames(obj.all) %in% colnames(obj.sub)]
```

## showing difference between obj and obj.sub

You can also embed plots, for example:

```{r pressure, echo=FALSE}
ST.FeaturePlot(obj, features = "RNA_clusters")+ST.FeaturePlot(obj.sub, features = "RNA_clusters")

FeatureOverlay(obj, features = "RNA_clusters")+FeatureOverlay(obj.sub, features = "lesion")

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
