---
title: "process sma objects"
author: "Jian Wu"
date: "2025-09-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load packages

```{r message=FALSE, warning=FALSE}
library(STutility)
library(ggplot2)
library(magrittr)
library(dplyr)
library(dbscan)
```

## load sma object data

obj.list.sma: object that is produced from multimodality integration, including both rna and msi modalities
obj.all.rna: object of rna that went through gene expression analysis and annotation

```{r pressure, echo=FALSE}

se.multi.list <- readRDS("/Users/wujian/Documents/MARCO/data/misc/se.multi.list")
se.RNA.list <- readRDS("/Users/wujian/Documents/MARCO/data/misc/se.RNA.list")
se.MSI.list <- readRDS("/Users/wujian/Documents/MARCO/data/misc/se.MSI.list")
se.merged.list <- readRDS("/Users/wujian/Documents/MARCO/data/misc/se.merged.list")

```


# plot aligned positions

```{r fig.width=12, fig.height=14}

aligned_coordinates <- do.call(rbind, lapply(names(se.merged.list), function(nm) {
  dims <- GetStaffli(se.merged.list[[nm]])@dims
  sf <- dims$`1`$width/dims$`2`$width
  coords <- cbind(se.merged.list[[nm]][[]], GetStaffli(se.merged.list[[nm]])@meta.data) |>
    tibble::rownames_to_column(var = "barcode") |>
    mutate(barcode = gsub(replacement = "_1", x = barcode, pattern = "_\\d")) |>
    select(barcode, warped_x, warped_y, nFeature_MSI) |>
    mutate(sample_type = case_when(is.na(nFeature_MSI) ~ "RNA",
                                   !is.na(nFeature_MSI) ~ "MSI")) |>
    mutate(warped_x = case_when(sample_type == "MSI" ~ warped_x*sf, TRUE ~ warped_x),
           warped_y = case_when(sample_type == "MSI" ~ warped_y*sf, TRUE ~ warped_y),
           sample_id = nm) |>
    select(barcode, warped_x, warped_y, sample_id, sample_type)
  return(coords)
}))

p <- ggplot(aligned_coordinates, aes(warped_x, warped_y,color = sample_type)) +
  geom_point(size = 0.5) +
  facet_wrap(~sample_id, ncol = 3)

p

```

# establish RNA-MSI mapping via nearest-neighbour classification

The RNA-MSI correspondance below is slightly different with the processing in the original article of SMA, i.e., the published code in GitHub by the authors
In M. Vicari et al's processing, a one-one correspondance between RNA-MSI is established, the shortcoming of which is that the RNA/Visium will suffer some spot lost.

What will be processed here is to avoid any removing spots in RNA/Visium objects due to RNA-MSI corresponding reason, and here is how this can be achieved.
The correspondance between the two modalities are not restricted to one-to-one type, but multiple RNA/Visium spots can correspond to the same MSI pixel, which is performed by: for each RNA/Visium spot, find its nearest-neighbor in MSI pixels (hence, the MSI pixels can not be garanteed to be used once, but rather, probably multiple times); the exact opposite way that is used in M. Vicari et al's processing.

```{r eval=FALSE}

# Split coordinates into a list
aligned_coordinates_list <- split(aligned_coordinates, aligned_coordinates$sample_id)

# Find the closest spatial neighbors 
knn_spatial_df_list <- lapply(aligned_coordinates_list, function(coords) {
  
  # get spatial coordinates for RNA and MSI data
  set1 = subset(coords, sample_type == "RNA")[, c("barcode", "warped_x", "warped_y")] |>
    data.frame(row.names = 1)
  set2 = subset(coords, sample_type == "MSI")[, c("barcode", "warped_x", "warped_y")] |>
    data.frame(row.names = 1)
  
  # Calculate pairwise distances using k nearest neighbors algorithm
  # here is the crucial difference from SMA's original program, where the positions of set1 and set2 have been swapped, and k has been changed to 1.
  knn_spatial_df <- kNN(x = set2, k = 1, query = set1)
  
  return(knn_spatial_df)
})


```

# integration

Assign the nearest neighbour in MSI pixels as the corresponding MSI assay to each RNA/Visium spot, where all RNA/Visium spots are preserved, at the cost of that some MSI pixels would be multiply reused.

```{r}

se.sma.list <- setNames(lapply(names(se.merged.list), function(nm) {

  knn_spatial_df <- knn_spatial_df_list[[nm]]

  se.multi <- se.RNA.list[[nm]]
  MSI <- se.MSI.list[[nm]]@assays$MSI@counts[, colnames(se.MSI.list[[nm]])[knn_spatial_df$id]]
  colnames(MSI) <- colnames(se.multi)
  se.multi[["MSI"]] <- CreateAssayObject(counts = MSI)

  return(se.multi)

}), nm = names(se.merged.list))

names(se.sma.list) <- names(se.merged.list)
```

# a visulization of the integration results between original SMA's and ours

original SMA's integration, named as se.multi.list
the integration used here, named as se.sma.list

```{r}

FeatureOverlay(se.multi.list[[1]], features = "nCount_MSI")+FeatureOverlay(se.sma.list[[1]], features = "nCount_MSI")

```

## the chuncks below are the same QC and joint analysis pipeline tha is used in the original SMA article

The original SMA object, i.e. se.multi.list, will be overwritten by se.sma.list, to facilitate going through the pipeline

```{r}
se.multi.list <- se.sma.list
```

## QC

```{r eval=FALSE}

se.multi.list <- lapply(se.multi.list, function(se.multi) {
  se.multi$nFeature_RNA <- colSums(GetAssayData(se.multi, slot = "counts", assay = "RNA") > 0)
  se.multi$nFeature_MSI <- colSums(GetAssayData(se.multi, slot = "counts", assay = "MSI") > 0)
  return(se.multi)
})

```
## Joint analysis

```{r eval=FALSE}

se.multi.list <- lapply(names(se.multi.list), function(nm) {
  
  se.multi <- se.multi.list[[nm]]
  cat(sprintf("Running analysis for sample %s", nm))
  
  # Normalize RNA data
  DefaultAssay(se.multi) <- 'RNA'
  se.multi <- NormalizeData(se.multi, verbose = FALSE) |> 
    FindVariableFeatures(verbose = FALSE) |> 
    ScaleData(verbose = FALSE) |> 
    RunPCA(verbose = FALSE) |>
    RunUMAP(reduction = "pca", dims = 1:30, reduction.name = "RNA.umap", verbose = FALSE) |>
    FindNeighbors(reduction = "pca", dims = 1:30, verbose = FALSE) |>
    FindClusters(verbose = FALSE)
  se.multi$RNA_clusters <- se.multi$seurat_clusters
  
  # Normalize MSI data
  DefaultAssay(se.multi) <- 'MSI'
  se.multi <- NormalizeData(se.multi, normalization.method = "CLR", verbose = FALSE) |> 
    FindVariableFeatures(nfeatures = 200, verbose = FALSE) |> 
    ScaleData(verbose = FALSE) |> 
    RunPCA(reduction.name = "msipca", dims = 1:10, verbose = FALSE) |>
    RunUMAP(reduction = "msipca", dims = 1:10, reduction.name = "MSI.umap", verbose = FALSE) |>
    FindNeighbors(reduction = "msipca", dims = 1:10, verbose = FALSE) |>
    FindClusters(verbose = FALSE)
  se.multi$MSI_clusters <- se.multi$seurat_clusters
  
  return(se.multi)
})

```

## Joint clustering

```{r}

se.multi.list <- lapply(se.multi.list, function(se.multi) {
  se.multi <- FindMultiModalNeighbors(
    se.multi, reduction.list = list("pca", "msipca"), 
    dims.list = list(1:30, 1:10), modality.weight.name = "RNA.weight"
  )
  return(se.multi)
})

```

## a weighted combination of RNA and protein data 

```{r eval=FALSE}

se.multi.list <- lapply(se.multi.list, function(se.multi) {
  se.multi <- RunUMAP(se.multi, nn.name = "weighted.nn", reduction.name = "wnn.umap", reduction.key = "wnnUMAP_", n.epochs = 1e3, min.dist = 0.15)
  se.multi <- FindClusters(se.multi, graph.name = "wsnn", algorithm = 3, resolution = 2, verbose = FALSE)
  se.multi$MM_clusters <- se.multi$seurat_clusters
  return(se.multi)
})

```

## swap names

```{r}
se.sma.list <- se.multi.list

se.multi.list <- readRDS("/Users/wujian/Documents/MARCO/data/misc/se.multi.list")

```

## display and save results

```{r}

FeatureOverlay(se.multi.list[[1]], features = "MM_clusters")+FeatureOverlay(se.sma.list[[1]], features = "MM_clusters")

```

```{r eval=FALSE, include=FALSE}

saveRDS(se.sma.list, file = "/Users/wujian/Documents/MARCO/R_objects/se.sma.list")

```

